Almost compatible functions and infinite length games

### 2016-09-06 articles/almost-compatible-functions-and-topological-games/article.tex
minor revisions from referee

            In other words, the family of subsets of any $X\in M$ in
            the extension $M[G]$ is equal to
             $\{ \tau[G] : \tau \subset X\times \mathbb P, \ \
            \tau\in M\ \}$.  We will be using
            the forcing poset is $\operatorname{Fn}(\omega_2, 2)$.
            The elements of this poset are all the finite partial functions
            from $\omega_2$ into $2$ ordered by reverse inclusion.
            It follows that, for any $\lambda\in \omega_2$,
             each of
             $\operatorname{Fn}(\lambda,2)$ and $\operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are subposets. For any
             $\operatorname{Fn}(\omega_2,2)$-generic
            filter  $G$, it easily follows that $G_\lambda = G\cap
            \operatorname{Fn}(\lambda,2)$ and
             $G^\lambda = G\cap
            \operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are also generic filters. But a much stronger statement is
             true.

@@

             Now let $Y$ equal $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda$. Furthermore,
             fix any $\mu\in \lambda\subset H$ such that
               $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\cap \lambda$ is contained in $\mu$.
             Let $\delta\in \omega_1$ denote the order type of
             $Y$ and let  $\varphi_{\mu,\lambda}$ be the order-preserving function
             from  $\mu\cup Y$ onto  the ordinal $\mu+\delta$.  This lifts
             canonically to
            an order-preserving bijection $\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)$.
            We can similarly make sense of the name
             $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$, call it $F_H$.
            Here simply, for each tuple $(\,(k,m), p)\in \dot
            f_{[\lambda,\lambda+\omega)}$,
             we have that $(\,(k,m),\varphi_{\mu,\lambda}(p))$ is in $F_H$.
            Again, let $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$
            be interpreted in the above sense as giving $F_H$ (which is an element
            of $H$).

@@

            \begin{proof}
            We will need the model constructed in \cite{MR1045371} in
             which an instance of Chang's conjecture
             $(\aleph_{\omega+1},\aleph_\omega) \rlap{$\rightarrow$}\,{\rightarrow}
             (\aleph_1,\aleph_0)$ is shown to fail.

@@

            The forcing notion $\mathbb P_0$ is simply the finite condition
            collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
             \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
             $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
             from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
             forcing extension we let $\mathbb P_1$ be the countable condition
             Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
             less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
             $\mathbb P_1$ has cardinality $\lambda$. We treat
            $\mathbb P*\mathbb P_1$ as containing $\mathbb P_0$
            as a subposet by identifying each $(p_0,1)$ with $p_0$.
            After forcing with $\mathbb P_0*\mathbb P_1$ we will have
            that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
             $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
             ordinal $\left(\lambda^{+\omega}\right)^V$.

            Now we assume that we have an assignment $\dot f_{\dot A}$ of a
             $\mathbb P_0*\mathbb P_1$-name of a finite-to-one function from
             $\dot A$ into $\omega$ for each
             $\mathbb P_0*\mathbb P_1$-name of a countable subset of
             $\lambda^{+\omega+1}$. We will obtain a contradiction.

@@

            Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
            Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
            (in that we have to formalize the enumeration of formulas as per
            G\"odel coding)
            let $\{\varphi_n : n\in \omega\}$ be an enumeration of all formulas
            in the language $L$, and let  $\ell_n$ be the minimal integer such
            that the free variables of $\varphi_n$ are among
             $\{ v_0, \ldots, v_{\ell_n}\}$.
            Then,   for each tuple
            $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
            of elements of $\lambda^{+\omega+1}$,
             we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
             the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
             $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
            If there is no such $\xi_0$,
               in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
            \ldots,\xi_{\ell_n})$, then set
             $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
            Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
             $X$
            of $N\cup \beta_0$ that satisfies that
             $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
             simply a large algebra, we can generate all the terms $t$ of
            the algebraic operations $\{ f_n : n\in \omega\}$.
            It is easily seen that for each $\zeta \in X$, there is a term
            $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
             for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
             each $\delta_i\in  N\cup \beta_0$.
             Assume that
              $\zeta\in \kappa^{+\omega+1}$. By
              re-indexing the variables in the term we can assume
              that there is an $n\leq m$ so that
               $\delta_i  <\beta_0$ for $1\leq i\leq n$
                and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
                Let $\vec a $ denote
                the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
              Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
             $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
             in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
             less naturally.
             Consider the set
               $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
            \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
             $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
             Now define $s_1(\eta,\vec a) $ to be
             $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
             and choose the
             unique $\zeta_1\in \lambda^{+\omega+1}$ such
             that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
             Note that $\alpha R_0\zeta_1$ holds if and only if
              $ \alpha \in s_0(\eta,\vec a)$, and therefore
             $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
             (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
             t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
             By elementarity then we have that $\zeta_1\in N$,
             and by similar reasoning the supremum, $\zeta_0$,
            of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
               This of course means that $\zeta < \xi_0$.
            \end{proof}

@@

  \begin{figure}[h]
\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=1em,minimum width=2em]
  {
    &
    \pl F \kmarkwin2\menGame{\oneptlind\kappa} &
    \pl F \kmarkwin2\schFillIntGame\kappa \\

@@

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\aleph_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\aleph_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\aleph_{\beta_{N+1}},C_{N+2}\cap\aleph_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\aleph_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning strategy and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\aleph_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\aleph_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\aleph_{\beta_{N+1}},C_{N+M+2}\cap\aleph_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning strategy.
  \end{proof}

@@

  \begin{proof}
    The proof proceeds nearly identically to the previous proof.
  \end{proof}
    % Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    % \(\schFillInitialGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    % for \(\pl F\) in \(\schFillInitialGame{\omega_\alpha}\) as follows:
    % \[
    %   \sigma(\<C\>,0)
    %     =
    %   \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    % \]
    % \[
    %   \sigma(\<C,D\>,n+1)
    %     =
    %   \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
    %     \cup
    %   \bigcup_{m\leq n}
    %   \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    % \]

>>>>

            In other words, the family of subsets of any $X\in M$ in
            the extension $M[G]$ is equal to
             $\{ \tau[G] : \tau \subset X\times \mathbb P, \ \
            \tau\in M\ \}$.  We will be using
            the forcing poset $\operatorname{Fn}(\omega_2, 2)$.
            The elements of this poset are all the finite partial functions
            from $\omega_2$ into $2$ ordered by reverse inclusion.
            It follows that, for any $\lambda\in \omega_2$,
             each of
             $\operatorname{Fn}(\lambda,2)$ and $\operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are subposets. For any
             $\operatorname{Fn}(\omega_2,2)$-generic
            filter  $G$, it easily follows that $G_\lambda = G\cap
            \operatorname{Fn}(\lambda,2)$ and
             $G^\lambda = G\cap
            \operatorname{Fn}(\omega_2\setminus
             \lambda,2)$ are also generic filters. But a much stronger statement is
             true.

@@

             Now let $Y$ equal $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\setminus \lambda$. Furthermore,
             fix any $\mu\in \lambda\subset H$ such that
               $\operatorname{supp}(\dot
             f_{[\lambda,\lambda+\omega)})\cap \lambda$ is contained in $\mu$.
             Let $\delta\in \omega_1$ denote the order type of
             $Y$ and let  $\varphi_{\mu,\lambda}$ be the order-preserving function
             from  $\mu\cup Y$ onto  the ordinal $\mu+\delta$.  This lifts
             canonically to
            an order-preserving bijection $\varphi_{\mu,\lambda}:
            \operatorname{Fn}(\mu\cup Y,2) \mapsto
            \operatorname{Fn}(\mu+\delta,2)$.
            We can similarly make sense of the name
             $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$, call it $F_H$.
            Here simply, for each tuple $(\,(\lambda+k,m), p)\in \dot
            f_{[\lambda,\lambda+\omega)}$,
             we have that $(\,(\mu+k,m),\varphi_{\mu,\lambda}(p))$ is in $F_H$.
            Again, let $\varphi_{\mu,\lambda}(\dot f_{[\lambda,\lambda+\omega)})$
            be interpreted in the above sense as giving $F_H$ (which is an element
            of $H$).

@@

            \begin{proof}
            We will need the model constructed in \cite{MR1045371} in
             which an instance of Chang's conjecture
             $(\aleph_{\omega+1},\aleph_\omega) \rlap{$\rightarrow$}\,{\rightarrow}
             (\aleph_1,\aleph_0)$ is shown to hold.

@@

            The forcing notion $\mathbb P_0$ is simply the finite condition
            collapse of $\kappa^{+\omega}$, i.e. $\mathbb P_0 = \left(
             \kappa^{+\omega}\right)^{<\omega}$. In the forcing extension by
             $\mathbb P_0$, one now has that the ordinal $\kappa^{+\omega+1}$
             from $V$ is the first uncountable cardinal $\aleph_1$. Then in this
             forcing extension we let $\mathbb P_1$ be the countable condition
             Levy collapse, $Lv(\lambda,\omega_2)$, which collapses all cardinals
             less than $\lambda$ to have cardinality at most $\aleph_1$. The poset
             $\mathbb P_1$ has cardinality $\lambda$. We treat
            $\mathbb P_0*\mathbb P_1$ as containing $\mathbb P_0$
            as a subposet by identifying each $(p_0,1)$ with $p_0$.
            After forcing with $\mathbb P_0*\mathbb P_1$ we will have
            that $\omega_1$ is the ordinal $\left(\kappa^{+\omega+1}\right)^V$,
             $\omega_2$ is the ordinal $\lambda$, and $\omega_\omega$ is the
             ordinal $\left(\lambda^{+\omega}\right)^V$.

            Now we assume that we have an assignment $\dot f_{\dot A}$ of a
             $\mathbb P_0*\mathbb P_1$-name of a finite-to-one function from
             $\dot A$ into $\omega$ for each
             $\mathbb P_0*\mathbb P_1$-name of a countable subset of
             $\lambda^{+\omega+1}$. We will obtain a contradiction
             to the claim of coherence.

@@

            Then set $\beta_0 = \sup(N\cap \kappa^{+\omega+1})$ and consider the
            Skolem closure $\mathop{Hull}(N\cup \beta_0, M)$.  A little informally
            (in that we have to formalize the enumeration of formulas as per
            G\"odel coding)
            let $\{\varphi_n : n\in \omega\}$ be an enumeration of all formulas
            in the language $L$, and let  $\ell_n$ be the minimal integer such
            that the free variables of $\varphi_n$ are among
             $\{ v_0, \ldots, v_{\ell_n}\}$.
            Then,   for each tuple
            $\langle \xi_1,\ldots, \xi_{\ell_n} \rangle$
            of elements of $\lambda^{+\omega+1}$,
             we define $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be
             the minimal $\xi_0\in \lambda^{+\omega+1}$ such that
             $M\models \varphi_n(\xi_0,\ldots, \xi_{\ell_n})$.
            If there is no such $\xi_0$,
               in other words if $M\models \lnot\exists x~~  \varphi_n(x,\xi_1,
            \ldots,\xi_{\ell_n})$, then set
             $f_n(\xi_1,\ldots, \xi_{\ell_n})$ to be $0$.
            Now $\mathop{Hull}(N\cup \beta_0,M)$ is just the minimal superset
             $X$
            of $N\cup \beta_0$ that satisfies that
             $f_n[X^{\{1,\ldots,\ell_n\}}] \subset X$ for all $n$. Since this is
             simply a large algebra, we can generate all the terms $t$ of
            the algebraic operations $\{ f_n : n\in \omega\}$.
            It is easily seen that for each $\zeta \in X$, there is a term
            $t(v_1,\ldots, v_m)$ such that $\zeta  = t(\delta_1,\ldots, \delta_m)$
             for some sequence $\langle \delta_1,\ldots, \delta_m\rangle$ with
             each $\delta_i\in  N\cup \beta_0$.
             Assume that
              $\zeta\in \kappa^{+\omega+1}$. By
              re-indexing the variables in the term we can assume
              that there is an $n\leq m$ so that
               $\delta_i  <\beta_0$ for $1\leq i\leq n$
                and $\kappa^{+\omega+1}\leq \delta_i$ for $n<i\leq m$.
                Let $\vec a $ denote
                the tuple $\langle \delta_{n+1},\ldots, \delta_m\rangle$.
              Choose $\eta\in N\cap \kappa^{+\omega+1}$ large enough so that
             $ \{  \delta_1, \ldots, \delta_n\}$ is contained in  $\eta$. Since set-membership
             in $M$ is coded by $R_0$ rather than $\in$ we have to argue a little
             less naturally.
             Consider the set
               $s_0(\eta,\vec a) = \{ t(\gamma_1,\ldots, \gamma_n,\vec a) :
            \{\gamma_1,\cdots,\gamma_n\} \in [\eta]^{\leq n}\}$. Clearly
             $s_0(\eta,\vec a)$ is a member of $H(\lambda^{+\omega+1})$.
             Now define $s_1(\eta,\vec a) $ to be
             $\{ x_\alpha : \alpha\in s_0(\eta, \vec a)\}$,
             and choose the
             unique $\zeta_1\in \lambda^{+\omega+1}$ such
             that $x_{\zeta_1} = s_1(\eta,\vec a)$. We claim that $\zeta_1\in N$.
             Note that $\alpha R_0\zeta_1$ holds if and only if
              $ \alpha \in s_0(\eta,\vec a)$, and therefore
             $$M\models (\forall \alpha)\left[\alpha R_0 \zeta_1 ~~\mbox{iff}~~
             (\exists   \gamma_1\in \eta)\cdots(\exists \gamma_n\in\eta)(\alpha=
             t(\gamma_1,\ldots, \gamma_n,\vec a))\right]~.$$
             By elementarity then we have that $\zeta_1\in N$,
             and by similar reasoning the supremum, $\zeta_0$,
            of $\zeta_1\cap \kappa^{+\omega+1}$ is also in $N$.
               This of course means that $\zeta < \beta_0$.
            \end{proof}

@@

  \begin{figure}[ht]
\begin{center}
\begin{tikzpicture}
  \matrix (m) [matrix of math nodes,row sep=3em,column sep=1em,minimum width=2em]
  {
    &
    \pl F \kmarkwin2\menGame{\oneptlind\kappa} &
    \pl F \kmarkwin2\schFillIntGame\kappa \\

@@

    Let \(\<C_0,C_1,\dots\>\) be an attack by
    \(\pl C\) in \(\schFillIntGame{\aleph_\alpha}\), and
    \(\alpha\in\bigcap_{n<\omega}C_n\).
    Choose \(N<\omega\) with \(\alpha<\aleph_{\beta_{N+1}}\). Consider the
    attack
    \(\<C_{N+1}\cap\aleph_{\beta_{N+1}},C_{N+2}\cap\aleph_{\beta_{N+1}},\dots\>\)
    by \(\pl C\) in \(\schFillIntGame{\aleph_{\beta_{N+1}}}\). Since
    \(\sigma_{N+1}\) is a winning \(2\)-mark and
    \(\alpha\in\bigcap_{n<\omega}C_{N+n+1}\cap\aleph_{\beta_{N+1}}\), either
    \(\alpha\in\sigma_{N+1}(\<C_{N+1}\cap\aleph_{\beta_{N+1}}\>,0)\) and thus
    \(\alpha\in\sigma(\<C_N,C_{N+1}\>,N+1)\), or
    \(
      \alpha
        \in
      \sigma_{N+1}(
        \<C_{N+M+1}\cap\aleph_{\beta_{N+1}},C_{N+M+2}\cap\aleph_{\beta_{N+1}}\>
        ,M+1
      )
    \)
    for some \(M<\omega\) and thus
    \(
      \alpha
        \in
      \sigma(
        \<C_{N+M+1},C_{N+M+2}\>,
        N+M+2
      )
    \). Thus \(\sigma\) is a winning \(2\)-mark.
  \end{proof}

@@

  \begin{proof}
    The proof proceeds nearly identically to the previous proof: substitute
    \(\alpha\in C_0\) in place of \(\alpha\in\bigcap_{n<\omega}C_n\) and proceed.
  \end{proof}
    % Let \(\sigma_n\) be a winning \(2\)-mark for \(\pl F\) in
    % \(\schFillInitialGame{\omega_{\beta_n}}\). Define the \(2\)-mark \(\sigma\)
    % for \(\pl F\) in \(\schFillInitialGame{\omega_\alpha}\) as follows:
    % \[
    %   \sigma(\<C\>,0)
    %     =
    %   \sigma_0(\<C\cap\omega_{\beta_0}\>,0)
    % \]
    % \[
    %   \sigma(\<C,D\>,n+1)
    %     =
    %   \sigma_{n+1}(\<D\cap\omega_{\beta_{n+1}}\>,0)
    %     \cup
    %   \bigcup_{m\leq n}
    %   \sigma_m(\<C\cap\omega_{\beta_m},D\cap\omega_{\beta_m}\>,n-m+1)
    % \]
